### 2017-05-18 Literature Review/A_Brief_Literature_Review.tex

\documentclass[12pt,preprint, authoryear]{elsarticle}

\usepackage{lmodern}
%%%% My spacing
\usepackage{setspace}
\setstretch{1.5}
\DeclareMathSizes{12}{14}{10}{10}

% Wrap around which gives all figures included the [H] command, or places it "here". This can be tedious to code in Rmarkdown.
\usepackage{float}
\let\origfigure\figure
\let\endorigfigure\endfigure
\renewenvironment{figure}[1][2] {
    \expandafter\origfigure\expandafter[H]
} {
    \endorigfigure
}

\let\origtable\table
\let\endorigtable\endtable
\renewenvironment{table}[1][2] {
    \expandafter\origtable\expandafter[H]
} {
    \endorigtable
}

\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
    \usepackage{xltxtra,xunicode}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi

\usepackage{amssymb, amsmath, amsthm, amsfonts}

\usepackage[round]{natbib}
\bibliographystyle{natbib}
\def\bibsection{\section*{References}} %%% Make "References" appear before bibliography
\usepackage{longtable}
\usepackage[margin=2cm,bottom=4cm,top=2.5cm, includefoot]{geometry}
\usepackage{fancyhdr}
\usepackage[bottom, hang, flushmargin]{footmisc}
\usepackage{graphicx}
\numberwithin{equation}{section}
\numberwithin{figure}{section}
\numberwithin{table}{section}
\setlength{\parindent}{0cm}
\setlength{\parskip}{1.3ex plus 0.5ex minus 0.3ex}
\usepackage{textcomp}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.3pt}

\usepackage{array}
\newcolumntype{x}[1]{>{\centering\arraybackslash\hspace{0pt}}p{#1}}

%%%%  Remove the "preprint submitted to" part. Don't worry about this either, it just looks better without it:
\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother

 \def\tightlist{} % This allows for subbullets!

\usepackage{hyperref}
\hypersetup{breaklinks=true,
            bookmarks=true,
            colorlinks=true,
            citecolor=blue,
            urlcolor=blue,
            linkcolor=blue,
            pdfborder={0 0 0}}

\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{5}

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}

%%% Include extra packages specified by user
% Insert custom packages here as follows
% \usepackage{tikz}

\begin{document}

\begin{frontmatter}  %

\title{Literature Review}

\author[Add1]{Fiona Ganie}
\ead{GNXFIO001@myuct.ac.za}

\address[Add1]{University of Cape Town}

\vspace{1cm}

\begin{keyword}
\footnotesize{
 \\ \vspace{0.3cm}
\textit{JEL classification} 
}
\end{keyword}
\vspace{0.5cm}
\end{frontmatter}

%________________________
% Header and Footers
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\pagestyle{fancy}
\chead{}
\rhead{}
\lfoot{}
\rfoot{\footnotesize Page \thepage\\}
\lhead{}
%\rfoot{\footnotesize Page \thepage\ } % "e.g. Page 2"
\cfoot{}

%\setlength\headheight{30pt}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%________________________

\headsep 35pt % So that header does not go over title

\section{\texorpdfstring{Introduction
\label{Introduction}}{Introduction }}\label{introduction}

Forecasting is a core element of the processes of a business. Producing
forecasts of high quality is essential as they play an important role in
inventory control, purchasing decisions of new equipment and investment
decisions, and are needed for effective planning (Granvik
\protect\hyperlink{ref-granvik2010forecasting}{2010}). Automatic
forecasting techniques have commonly been used in business to produce a
wide variety of forecasts and are a useful tool for producing a large
number of forecasts. Automatic techniques can be used by non-experts
without training in the use of time series (R. J. Hyndman, Khandakar,
and others \protect\hyperlink{ref-hyndman2007automatic}{2007}).
Automatic forecasting algorithms must have the ability to select the
appropriate model for forecasting and estimate model parameters, without
the need for human intervention (R. J. Hyndman, Khandakar, and others
\protect\hyperlink{ref-hyndman2007automatic}{2007}). The most common
automatic forecasting techniques used are the ARIMA and Exponential
Smoothing (ETS) models. ETS models have been used extensively in
forecasting demand for inventories (Billah et al.
\protect\hyperlink{ref-billah2006exponential}{2006}). They have
performed well in forecasting competitions against more complex models
and are particularly good at forecasting seasonal data over short
horizons (R. J. Hyndman et al.
\protect\hyperlink{ref-hyndman2002state}{2002}).

Automatic forecasting techniques can be too inflexible and do not allow
for analysts with specialized knowledge to incorporate their expertise
into the model effectively. This leads to difficulties in producing
forecasts of high quality, thus the demand for high quality forecasts
far exceeds the rate at which they are produced (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). In December 2016,
Facebook released their forecasting model, Prophet. According to Taylor
and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017}),
Prophet is a flexible model that can be adjusted by a large number of
non-experts who have little knowledge about time series, however have
expert knowledge on the data-generating process. Prophet allows for a
large number of forecasts to be produced across a variety of problems
and consists of a robust evaluation system that allows for a large
number of forecasts be evaluated and compared. This is Facebook's
definition of forecasting at scale, and provides a solution to the
problems that automatic forecasting pose (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\section{\texorpdfstring{Exponential State Space Models
\label{Exponential State Space Models}}{Exponential State Space Models }}\label{exponential-state-space-models}

\subsection{Exponential Smoothing
Methods}\label{exponential-smoothing-methods}

ETS methods originated in the 1950's and is the key work of Brown ,
Holt, and Winters (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). It has been widely used for
many applications in industry and has been the foundation of many
successful forecasting methods (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). Exponential smoothing
techniques produce forecasts by calculating them as a weighted average
of past observations. Observations that are most recent are more heavily
weighted, with these weights decaying exponentially as the observations
get older.

Exponential smoothing methods were originally classified by Pegel
according to the components present in the time series (trend and
seasonality), and whether they were additive or multiplicative in nature
(De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). Gardner
and Taylor further extended this classification to include a damped
additive and multiplicative trend respectively (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). A taxonomy of exponential
smoothing methods was later created by Hyndman et al. and extended by
Taylor (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
It considered all possible combinations of the trend and seasonality
components thus producing 15 methods of exponential smoothing as seen in
table \ref{tab1}.

\begin{table}[H]
\centering
\begin{tabular}{rlll}
  \hline
 & Trend &  & Seasonality \\ 
  \hline
 & None(N) & Additive(A) & Multiplicative(M) \\ 
  None(N) & (N,N) & (N,A) & (N,M) \\ 
  Additive(A) & (A,N) & (A,A) & (A,M) \\ 
  Additive\_damped(Ad) & (Ad,N) & (Ad,A) & (Ad,M) \\ 
  Multiplicative(M) & (M,N) & (M,A) & (M,M) \\ 
  Multiplicative\_damped(Md) & (Md,N) & (Md,A) & (Md,M) \\ 
   \hline
\end{tabular}
\caption{Taxonomy of Exponential Smoothing Methods \label{tab1}} 
\end{table}

\hfill

The most common of these methods are Simple Exponential Smoothing (N,N),
Holt's linear method (A,N) and Holt-Winters seasonal method (A,A).

Although exponential smoothing methods performed well and have been
widely used in business, they did not have any statistical framework
underpinning them (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). They produced point estimates
rather than prediction intervals and selection of a forecasting method
was specific to the time series in question. Efforts were made by Box
and Jenkins to create a statistical framework that underlies exponential
smoothing methods and they found that forecasts produced by linear
methods are special cases of ARIMA models (De Gooijer and Hyndman
\protect\hyperlink{ref-de200625}{2006}). This however does not hold for
the multiplicative methods.

\subsection{State Space Models}\label{state-space-models}

The motivation behind state space models was the need for a statistical
framework that underlies all methods of exponential smoothing (De
Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}). R. J.
Hyndman et al. (\protect\hyperlink{ref-hyndman2002state}{2002}) provided
this framework as an extension of Ord, Koehler, and Snyder's class of
exponential state space models which underpinned some of the smoothing
methods (De Gooijer and Hyndman \protect\hyperlink{ref-de200625}{2006}).
He proposed two state space models, one corresponding to an additive
error and the other to a multiplicative error. By specifying a
distribution for these error terms, he turned the previously
deterministic methods into stochastic models. These two models were then
applied to all 15 smoothing methods, producing 30 models in total.

The two state space models produce identical point estimates to each
other as well as to the exponential smoothing methods, however they will
generate differing prediction intervals (R. Hyndman and Athanasopoulos
\protect\hyperlink{ref-hyndman2013forecasting}{2013}). Each model
consists of a measurement equation and a transition equation. The
measurement equation shows the relationship between the observations and
the unobserved components (level, trend seasonality), while the
transition equation shows how the unobserved components change over
time.

The introduction of the exponential state space model has allowed for
model selection using information criteria rather than using an ad hoc
method. It has also made the calculation of the likelihood easier and
simulation from the underlying model possible (R. J. Hyndman et al.
\protect\hyperlink{ref-hyndman2002state}{2002}).

\subsection{Application and
Performance}\label{application-and-performance}

In a paper by Smith and Agrawal
(\protect\hyperlink{ref-smith2015comparison}{2015}), Holt-Winters
exponential smoothing (HWES) model was applied to time series data
relating to data on patents. The patent data was classified into three
different groups; monthly forecasts were made and the results were
compared to forecasts produced by ARIMA models. To avoid
over-forecasting, the trend of the HWES was damped to produce forecasts
with a flat trend for later observations. The results showed that all
models forecasted the data adequately. The HWES model fitted the data
well and performed better than the ARIMA models in the case where the
time series for the patent group was fairly stationary. Based on the
forecasting results, it could not be concluded that a specific time
series model is better for forecasting patent data.

Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
forecasted tourism demand in specific European countries over a short,
medium and long term horizon. The ETS model was used to produce
forecasts and was compared to six parametric and non-parametric
techniques. The results showed that no single model is best for
producing forecasts across all countries and time horizons. The ETS
model together with the Neural Networks (NN) and Fractionalized ARIMA
(ARFIMA) produced forecasts with the least amount of accuracy and are
thus not relevant models for forecasting tourism demand in Europe. 24
steps-ahead error forecasts produced by the ETS model were found to be
larger than the errors produced by the benchmark model - Recurrent
Single Spectrum Analysis (SSA-R) across all European countries.

\subsection{Robustness}\label{robustness}

Gardner (\protect\hyperlink{ref-gardner2006exponential}{2006}) explained
the robustness of ETS models by comparing them to equivalent models.
Simple exponential smoothing which is equivalent to an ARIMA(0,1,1)
model is the most robust method and has performed well in forecasting
many types of time series. By comparing simple exponential smoothing to
lower order ARIMA models, it is seen that errors that arise due to model
specification is not as much of a problem as ARIMA models. Furthermore,
ARIMA models tend to produce larger MSE due to model selection errors,
and this is worsened if the errors are not normally distributed (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

Satchell and Timmermann found that for time series with a finite
history, the weights assigned to each observation in the simple
exponential smoothing method are robust, provided that the variance of
the random walk relative to the variance of the error component is not
too small (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

Forecasts produced by simple exponential smoothing have performed well
in modelling annual sales of different products, as well as aggregated
economic series with low sampling frequencies (Gardner
\protect\hyperlink{ref-gardner2006exponential}{2006}).

\section{Prophet}\label{prophet}

\subsection{Bayesian Generalized Additive
Model}\label{bayesian-generalized-additive-model}

Prophet is the key work of Taylor and Letham
(\protect\hyperlink{ref-taylor2017forecasting}{2017}) and consists of a
decomposable model with a component for growth, seasonality, and
holidays. These components consist of linear and non-linear functions of
time. This is similar to a Generalized Additive Model (GAM) which is a
regression model that consists of non-linear and linear regression
functions applied to predictor variables (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Prophet, like GAM,
frames the forecasting problem as a curve fitting exercise using
backfitting to find the regression functions.

The growth component is modelled in a similar way to modelling
population growths which uses a logistic growth model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Populations
typically grow non-linearly (although the growth component could also be
linear) up to an upper bound known as the carrying capacity, and remains
constant thereafter. The rate at which the population grows changes over
time, and this is accounted for by including changepoints in the model
where the growth rate may be changed (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). These changepoints
can be automatically selected and have a Laplace prior distribution
placed on them (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). The parameter of
the prior distribution can be used to adjust the growth rate of the
changepoints. This allows non-experts with knowledge about events that
may affect growth to use the parameter as a knob and adjust it to
increase or decrease the number of changepoints (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). It also allows for
the analyst to add changepoints which the automatic selection procedure
may have missed (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). Furthermore,
analysts may also specify the carrying capacity and adjust it based on
their knowledge of the total market size (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

The decomposable form of the model allows for components to be easily
added to it. This allows for multiple seasonality components with
different periods to be added to the model. The seasonality component is
modelled by a Fourier series, with the parameters of the Fourier series
having a Normal prior distribution (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). The variance of
the parameters can be adjusted by analysts to smooth the model and
change how much of historical seasonality is projected to the future
(Taylor and Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}).

The name, date, and country of past and future holidays and events may
be inputted by the analyst into a list (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}). By specifying the
country in which the events take place or the holidays occur, separate
lists can be populated for global events/holidays and country-specific
events/holidays. The union of the two lists are then used for
forecasting. Like seasonality, a Normal prior distribution is placed on
the parameters of the holidays, and the variance of the parameters can
be adjusted by analysts to smooth the model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\subsection{Application and
Performance}\label{application-and-performance-1}

Taylor and Letham (\protect\hyperlink{ref-taylor2017forecasting}{2017})
forecasted the number of events on Facebook using Prophet. The time
series was impacted by holidays, had strong multi-period seasonality,
and a piecewise trend. The forecasts produced by Prophet were compared
to forecasts produced by common automatic forecasting techniques such as
ETS, ARIMA, and the seasonal naïve model, as well as to simple models
such as the naïve model. While the ETS and seasonal naïve model were
quite robust, the ARIMA forecasts were fragile. No model besides Prophet
accounted for the dips around holidays and the upward trend of the time
series towards later observations. Hence, Prophet had lower forecasting
errors compared to the automatic forecasting methods.

\subsection{Robustness}\label{robustness-1}

Prophet is a flexible model and has intuitive parameters that can be
easily interpreted by human beings. The GAM is fitted quickly, allowing
the analyst to interactively change the model parameters (Taylor and
Letham \protect\hyperlink{ref-taylor2017forecasting}{2017}). Unlike
ARIMA models, Prophet can produce forecasts over different scales and
allows for missing values in the time series without the need for
interpolation (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

When a large number of forecasts are produced, manually identifying
problematic forecasts becomes a time consuming and difficult task.
Prophet provides a semi-automated forecast evaluation system that
selects the best model that fits the data. When there are large forecast
errors, the forecasts are flagged so that the analyst can explore the
cause of the errors, identify and remove potential outliers and either
adjust the model or choose a more appropriate model (Taylor and Letham
\protect\hyperlink{ref-taylor2017forecasting}{2017}).

\section{Evaluation of Forecasts}\label{evaluation-of-forecasts}

\subsection{Common Methods}\label{common-methods}

Model selection has historically been very subjective, with no
techniques that offered mathematical rigour for choosing models (Smith
and Agrawal \protect\hyperlink{ref-smith2015comparison}{2015}). Progress
has been made over time and many mathematical techniques have been
introduced for model selection. Use of information criteria is one such
method, with Akaike's information criterion (AIC) and Bayes information
criterion (BIC) being the most common (Smith and Agrawal
\protect\hyperlink{ref-smith2015comparison}{2015}). These information
criteria minimize forecasting errors while penalizing it for
overfitting. Information criteria should be used to compare models of
similar structure only. To compare models of different structure, the
most commonly used measures of forecast errors are the Root Mean Squared
error (RMSE), Mean Absolute Error (MAE), Mean Absolute Percentage Error
(MAPE), and Mean Absolute Scaled Error (MASE). A lower value of these
statistics indicates that the forecasts are better, with a MASE less
than one indicating that the forecasts are better than those produced by
a naïve model (Smith and Agrawal
\protect\hyperlink{ref-smith2015comparison}{2015}).

Smith and Agrawal (\protect\hyperlink{ref-smith2015comparison}{2015})
have used these methods in their paper when evaluating patent forecasts
and making comparisons between the HWES and ARIMA models. In a paper by
Hassani et al. (\protect\hyperlink{ref-hassani2015forecasting}{2015})
they chose their ETS model based on the selection that R automatically
made for the time series. They compared models of different structure
using the RMSE and a modified Diebold-Mariano test. R. J. Hyndman et al.
(\protect\hyperlink{ref-hyndman2002state}{2002}) suggests that ETS
models should be selected using the AIC rather than the MSE or MAPE. The
AIC provides a way of choosing between models with an additive error and
models with a multiplicative error since it is based on the likelihood
function. The MSE and MAPE are not able to select between the two error
types because both models produce the same point estimates. According to
De Gooijer and Hyndman (\protect\hyperlink{ref-de200625}{2006}), the
best approach of selecting models with different structure is the
Diebold-Mariano test. The MSE is scale dependent and should not be used
to make comparisons while the MAPE encounters difficulties when the time
series values are close to or equal to zero.

\subsection{Mincer-Zarnowitz Approach}\label{mincer-zarnowitz-approach}

Mincer and Zarnowitz
(\protect\hyperlink{ref-mincer1969evaluation}{1969}) proposed two
methods of measuring forecast accuracy - an absolute and a relative
measure. Absolute measures of accuracy measure the distance between
actual and predicted values. If the mean distance is equal to zero, the
forecast is said to be unbiased (Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). To analyse absolute
forecast errors, the observed values are regressed against the predicted
values. A joint hypothesis test is performed to check if the intercept
is equal to zero and the slope is equal to one. If the intercept is
equal to zero, the forecasts are unbiased and do not overestimate or
underestimate the data. If the slope is equal to one, the residual
errors are uncorrelated with the predictions and the forecast is said to
be efficient (Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). The MSE is a
measure that can be decomposed into the residual variance and two
components, one reflecting the bias and the other reflecting the
inefficiency of the forecast. If the forecasts are unbiased and
efficient, the MSE will reduce to the residual variance (Mincer and
Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}). Absolute
forecast measures cannot be used to make comparisons between forecasts
with different scales or economic variables. Thiel states that the
consequences of forecasting errors and how they impact decisions is of
greater importance than the size of the forecasting errors (Mincer and
Zarnowitz \protect\hyperlink{ref-mincer1969evaluation}{1969}).

Rather than using absolute measures of forecast accuracy, Mincer and
Zarnowitz (\protect\hyperlink{ref-mincer1969evaluation}{1969}) suggested
the use of relative measures of forecast accuracy. Relative accuracy
analysis allows for meaningful comparisons of different forecasting
methods to be made. They proposed an index which considers the ratio of
the MSE of the forecast to the MSE of a benchmark forecasting method.
They used an extrapolation of the data history as a benchmark, as it is
a cost effective and accessible method, however the benchmark may be any
method which is relevant for comparison. This ratio is known as the
Relative Mean Square Error (RM).

The numerator represents a return which is inversely proportional to the
MSE error of the forecasts, while the denominator represents the cost of
forecasting which is inversely proportional to the MSE of the benchmark
(Mincer and Zarnowitz
\protect\hyperlink{ref-mincer1969evaluation}{1969}). Hence the ratio is
representative of a rate of return index and ranks the performance of
forecasts as such. If the forecasts are better than the benchmark, the
RM will be less than one.

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-billah2006exponential}{}
Billah, Baki, Maxwell L King, Ralph D Snyder, and Anne B Koehler. 2006.
``Exponential Smoothing Model Selection for Forecasting.''
\emph{International Journal of Forecasting} 22 (2). Elsevier: 239--47.

\hypertarget{ref-de200625}{}
De Gooijer, Jan G, and Rob J Hyndman. 2006. ``25 Years of Time Series
Forecasting.'' \emph{International Journal of Forecasting} 22 (3).
Elsevier: 443--73.

\hypertarget{ref-gardner2006exponential}{}
Gardner, Everette S. 2006. ``Exponential Smoothing: The State of the
Art---Part Ii.'' \emph{International Journal of Forecasting} 22 (4).
Elsevier: 637--66.

\hypertarget{ref-granvik2010forecasting}{}
Granvik, Ashley. 2010. ``Forecasting Exchange Rates.'' Arcada-Nylands
svenska yrkeshögskola.

\hypertarget{ref-hassani2015forecasting}{}
Hassani, Hossein, Emmanuel Sirimal Silva, Nikolaos Antonakakis, George
Filis, Rangan Gupta, and others. 2015. ``Forecasting Accuracy Evaluation
of Tourist Arrivals: Evidence from Parametric and Non-Parametric
Techniques.''

\hypertarget{ref-hyndman2013forecasting}{}
Hyndman, RJ, and G Athanasopoulos. 2013. ``Forecasting: Principles and
Practice, Otexts, Australia.'' \emph{Available via: Www. OTexts.
Com/Fpp}.

\hypertarget{ref-hyndman2007automatic}{}
Hyndman, Rob J, Yeasmin Khandakar, and others. 2007. ``Automatic Time
Series for Forecasting: The Forecast Package for R.'' Monash University,
Department of Econometrics; Business Statistics.

\hypertarget{ref-hyndman2002state}{}
Hyndman, Rob J, Anne B Koehler, Ralph D Snyder, and Simone Grose. 2002.
``A State Space Framework for Automatic Forecasting Using Exponential
Smoothing Methods.'' \emph{International Journal of Forecasting} 18 (3).
Elsevier: 439--54.

\hypertarget{ref-mincer1969evaluation}{}
Mincer, Jacob A, and Victor Zarnowitz. 1969. ``The Evaluation of
Economic Forecasts.'' In \emph{Economic Forecasts and Expectations:
Analysis of Forecasting Behavior and Performance}, 3--46. NBER.

\hypertarget{ref-smith2015comparison}{}
Smith, Mick, and Rajeev Agrawal. 2015. ``A Comparison of Time Series
Model Forecasting Methods on Patent Groups.'' In \emph{MAICS}, 167--73.

\hypertarget{ref-taylor2017forecasting}{}
Taylor, Sean J, and Benjamin Letham. 2017. ``Forecasting at Scale.''

% Force include bibliography in my chosen format:
\newpage
\nocite{*}
\bibliography{}

\end{document}